---
title: "Data Import"
output: github_document
---
```{r setup, echo = FALSE, message= FALSE}
library(tidyverse)
library(readxl)
library(haven)
```


This document will show how to import data.

## Import the FAS Litters CSV and Fixing Column Names 
```{r}
litters_df = read_csv("data/FAS_litters.csv") 
litters_df = janitor::clean_names(litters_df)
```
Only using the "clean names" function inside janitor package so we aren't going to load whole janitor package


## Look at the dataset
```{r}
litters_df

head(litters_df)
tail(litters_df, 10)
```

```{r eval= FALSE}
view(litters_df)
```
Want this to be False when using view bc when you knit it it gets confused

## Learning Assessement Load FAS_Pups.CSV
use relative paths
```{r}
pups_df = read_csv("data/FAS_pups.csv")
pups_df = janitor::clean_names(pups_df)
pups_df
```

use absolute path is putting whole path of file (couldnt make it work in class I think I fixed it)
```{r eval=FALSE}
pups_df=read_csv(("/Users/meredithsteinberg/Desktop/data_wrangling_i/data/"))
```

## Look at read csv options

col_names and skipping rows

```{r}
litters_df=
  read_csv(
    file="data/FAS_litters.csv",
    col_names=FALSE,
    skip=1
  )
```
skipping line bc of excel file data starts at second line

What about missing data
```{r}
litters_df=
  read_csv(
    file="data/FAS_litters.csv",
    na = c("NA", "", ".")
  )
view(litters_df)
```
listing three things in c function that are missing values so they all read NA 

Looking at a specific column
```{r}
litters_df=janitor::clean_names((litters_df))
pull(litters_df, gd0_weight)
```

What if we code 'group' as a factor variable?
```{r}
litters_df=
  read_csv(
    file= "data/FAS_litters.csv",
    na = c("NA", "", "."),
    col_types = cols(
      Group= col_factor()
    )
  )
```


## Import excel file

import MLB 2011 summary data

```{r}
mlb_df = read_excel("data/mlb11.xlsx", sheet="mlb11") 
mlb_df
```

## Import SAS data

```{r}
pulse_df=read_sas("data/public_pulse_data.sas7bdat")
pulse_df
```

## Never use read.csv() 

```{r}
litters_df= read.csv("data/FAS_litters.csv")
litters_df
```

Never do this either bc we are taking a variable out of a dataframe
```{r}
litters_df$L
```



