---
title: "Data Manipulation"
output: github_document
---
```{r setup, echo = FALSE, message= FALSE}
library(tidyverse)
```


This document will show how to *manipulate* data.

Import the two datasets that we're going to manipulate.


```{r}
litters_df = read_csv("data/FAS_litters.csv" , na= c("NA", "", "."))
litters_df = janitor::clean_names(litters_df)
litters_df

pups_df=
  read_csv("data/FAS_pups.csv", na = c("NA", "", "."))
pups_df = janitor::clean_names(pups_df)
pups_df
```

##'select'
Use 'select' to select variables

```{r}
select(litters_df, group, litter_number, gd0_weight)
```

```{r}
select(litters_df, group:gd18_weight)
```
giving columbs from from group to gd18 weight

```{r}
select(litters_df, -pups_survive)
```
getting rid of pups_survive

```{r}
select(litters_df, -(group:gd18_weight))
```

```{r}
select(litters_df, starts_with("gd"))
```
only gives us columns that start w gf

```{r}
select(litters_df, contains("pups"))
```
only gives us columns that includes pups

```{r}
select(litters_df, GROUP=group)
```
renaming the variable group

```{r}
rename(litters_df, GROUP=group)
```
another way to rename one variable

```{r}
select(litters_df, litter_number, gd0_weight, everything())
```
if wanted to move things around and put in order we want (litter number and gd0 weight) and keep everything else same

```{r}
relocate(litters_df, litter_number, gd0_weight)
```
dont have to say everything in this version will just have these two variables first

### learning assessment for select-- pups dataset

```{r}
select(pups_df, litter_number, sex, pd_ears)
```



## 'filter'


```{r}
filter(litters_df, gd_of_birth ==20)
filter(litters_df, gd_of_birth ==19)

```
double equal is asking if it is equal to 20 or 19. will give us filter values we are asking for if they exist

```{r}
filter(litters_df, pups_born_alive >5)
filter(litters_df, pups_born_alive >= 8)

```

```{r}
filter(litters_df, pups_born_alive != 9)
```
filtering out rows NOT equal to 9 for pups born alive


```{r}
filter(litters_df, group == "Con7")
filter(litters_df, group %in% c("Low7","Low8"))
```
first line only seeing group that has Con7
Low 7 or Low 8 in second line


```{r}
filter(litters_df, group %in% c("Low7","Low8"), pups_born_alive ==8)
```
will give us only datapoints with group Low7 or Low8 value and pups born alive equal to 8

```{r}
drop_na(litters_df)

drop_na(litters_df, gd0_weight)

```
get rid of missing data. can drop NAs in everything or can drop NAs for specific variables

###learning assessment pups
```{r}
filter(pups_df, sex =="1")
filter(pups_df, pd_walk < "11" , sex =="2")
```


## 'mutate'

```{r}
mutate(litters_df, wt_gain = gd18_weight - gd0_weight)

mutate(litters_df, sq_pups = pups_born_alive^2)

mutate(litters_df, group = str_to_lower(group))
```
first two lines creating new variable
last line lowercasing all group variables

```{r}
mutate(
  litters_df,
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group)
  )
```
putting it all together


## 'arrange'

```{r}
arrange(litters_df, gd0_weight)
arrange(litters_df, desc(gd0_weight))
```
order of increasing or descending gd0 weight

```{r}
arrange(litters_df, pups_born_alive, gd0_weight)
```
order increasing by pups born alive and then gd0 weight



## 'PIPING'

